https://www.jstatsoft.org/article/view/v052i04
lgcp package doesn't work as because the suplimentary matterial doesn't work and gives the same error as the code is was trying
```{r}
# Load required libraries
library(lgcp)
library(spatstat.geom)
library(dplyr)
library(readxl)

# Set working directory and load window data
setwd("C:/Users/ajmoo/OneDrive - Queen's University Belfast/SOR4001 Project/R/DATA")

# Load the window data (boundary coordinates for each time point)
windowcontrol <- read.table("control_x3.txt", sep = "\t")
windowcontrol <- owin(poly = as.matrix(windowcontrol))
window3d <- read.table("3d_pONT_x1.txt", sep = "\t")
window3d <- owin(poly = as.matrix(window3d))
window7d <- read.table("7d_pONT_x1.txt", sep = "\t")
window7d <- owin(poly = as.matrix(window7d))

# Load and prepare point data
PointControl <- read_excel("control_0d_Points.xlsx")
Point3d <- read_excel("3d_pONT_Points.xlsx")
Point7d <- read_excel("7d_pONT_Points.xlsx")
Point_df_control <- as.data.frame(PointControl)
Point_df_3d <- as.data.frame(Point3d)
Point_df_7d <- as.data.frame(Point7d)

# Convert area from pixels to meters squared
Point_df_control$Area_meters_squared <- Point_df_control$Area1 * (1.38 * 0.001)^2
Point_df_3d$Area_meters_squared <- Point_df_3d$Area1 * (1.38 * 0.001)^2
Point_df_7d$Area_meters_squared <- Point_df_7d$Area1 * (1.38 * 0.001)^2

# Define circle parameters (same for all windows)
circle_center <- c(x = 2500, y = 4000)  # Replace with actual coordinates
circle_radius <- 100  # Define radius in the same units as the data

# Function to filter points within the circle
filter_points_in_circle <- function(points_df, center, radius) {
  points_df %>%
    filter((X1 - center[1])^2 + (Y1 - center[2])^2 <= radius^2)
}

# Filter points for each dataset
Point_circle_control <- filter_points_in_circle(Point_df_control, circle_center, circle_radius)
Point_circle_3d <- filter_points_in_circle(Point_df_3d, circle_center, circle_radius)
Point_circle_7d <- filter_points_in_circle(Point_df_7d, circle_center, circle_radius)

# Optional: Randomly sample up to 500 points from the filtered data
Point_sample_control <- Point_circle_control %>% sample_n(min(n(), 500))
Point_sample_3d <- Point_circle_3d %>% sample_n(min(n(), 500))
Point_sample_7d <- Point_circle_7d %>% sample_n(min(n(), 500))

# Add a time column to each dataset
Point_sample_control$time <- 0  # Control corresponds to t = 0
Point_sample_3d$time <- 3  # 3-day dataset corresponds to t = 3
Point_sample_7d$time <- 7  # 7-day dataset corresponds to t = 7

# Summary of results
cat("Control: Points in circle =", nrow(Point_circle_control), "\n")
cat("3d: Points in circle =", nrow(Point_circle_3d), "\n")
cat("7d: Points in circle =", nrow(Point_circle_7d), "\n")

Point_sample_3d
```

```{r}
# Function to plot a circle on the window
plot_circle_on_window <- function(window, circle_center, circle_radius, title) {
  # Create the circle as a polygon
  theta <- seq(0, 2 * pi, length.out = 200)  # Generate points for a smooth circle
  circle_x <- circle_center[1] + circle_radius * cos(theta)
  circle_y <- circle_center[2] + circle_radius * sin(theta)
  
  # Plot the window
  plot(window, main = title, col = "lightblue", border = "blue")
  
  # Add the circle to the plot
  lines(circle_x, circle_y, col = "red", lwd = 2)
  
  # Indicate the circle's center
  points(circle_center[1], circle_center[2], col = "black", pch = 16)
}

# Plot the circle on each window
par(mfrow = c(1, 3))  # Arrange plots in a single row
plot_circle_on_window(windowcontrol, circle_center, circle_radius, "Control Window")
plot_circle_on_window(window3d, circle_center, circle_radius, "3d Window")
plot_circle_on_window(window7d, circle_center, circle_radius, "7d Window")

```



*rLGCP*
*Try this next, its a LCGP using spats stats tutorial*
https://towardsdatascience.com/fitting-point-process-models-in-r-263db85e12dc

LGCP with no time parameter
```{r}
# Load required libraries
library(spatstat)
library(sf)
library(sp)
library(raster)
library(fields)
library(rstan)
library(tidyverse)
library(bayesplot)
library(dplyr)

# Function to simulate and analyze LGCP using your data
analyze_lgcp <- function(points_df, center, radius, b0, b1, var, scale, plotdat = TRUE) {
  # Define circular window
  win <- owin(
    xrange = c(center[1] - radius, center[1] + radius),
    yrange = c(center[2] - radius, center[2] + radius)
  )
  win <- intersect.owin(win, disc(radius = radius, centre = center)) # Ensure circular window

  # Convert sampled points to ppp object
  points_ppp <- ppp(
    x = points_df$X1,
    y = points_df$Y1,
    window = win
  )
  
  # Define the grid for environmental covariate (angle calculation)
  x_range <- seq(win$xrange[1], win$xrange[2], length.out = 100)  # Increase grid resolution
  y_range <- seq(win$yrange[1], win$yrange[2], length.out = 100)  # Increase grid resolution
  
  # Calculate angle around the circle for each point
  gridcov <- outer(x_range, y_range, function(x, y) {
    # Calculate angle in radians
    atan2(y - center[2], x - center[1])
  })

  # Convert covariate to im object
  cov_im <- im(gridcov, xcol = x_range, yrow = y_range)

  # Simulate the LGCP
  lgcp_sim <- rLGCP(
    model = "exp",
    mu = im(b0 + b1 * gridcov, xcol = x_range, yrow = y_range),
    var = var,
    scale = scale,
    win = win
  )
  
  # Plot results if requested
  if (plotdat) {
    par(mfrow = c(1, 2), mar = c(2, 2, 1, 1))
    plot(points_ppp, main = "Observed Point Pattern")
    plot(lgcp_sim, main = "Simulated LGCP")
  }

  # Return simulated LGCP and observed points for further analysis
  list(simulated = lgcp_sim, observed = points_ppp)
}

# Parameters
b0 <- -5.5   # Intercept
b1 <- 0.2   # Coefficient for covariate (angle)
var <- 0.1    # Variance
scale <- 50  # Scale'

# Plot observed points with time as color
plot(result_control$observed, main = "Observed Points - Control")
plot(result_3d$observed, main = "Observed Points - 3 Day")
plot(result_7d$observed, main = "Observed Points - 7 Day")

# Analyze each dataset
cat("Analyzing Control dataset...\n")
result_control <- analyze_lgcp(Point_sample_control, circle_center, circle_radius, b0, b1, var, scale)

cat("Analyzing 3-day dataset...\n")
result_3d <- analyze_lgcp(Point_sample_3d, circle_center, circle_radius, b0, b1, var, scale)

cat("Analyzing 7-day dataset...\n")
result_7d <- analyze_lgcp(Point_sample_7d, circle_center, circle_radius, b0, b1, var, scale)

summary(result_control$simulated) 
summary(result_control$observed)#set b0 equal to log(avg intensity) = -5.4

summary(result_3d$simulated)
summary(result_3d$observed)

summary(result_7d$simulated)
summary(result_7d$observed)



```
Incorporating Time as a Covariate
```{r}
# Load required libraries
library(spatstat)
library(sf)
library(sp)
library(raster)
library(fields)
library(rstan)
library(tidyverse)
library(bayesplot)
library(dplyr)

# Function to simulate and analyze LGCP using your data
analyze_lgcp <- function(points_df, center, radius, b0, b1, var, scale, plotdat = TRUE) {
  # Define circular window
  win <- owin(
    xrange = c(center[1] - radius, center[1] + radius),
    yrange = c(center[2] - radius, center[2] + radius)
  )
  win <- intersect.owin(win, disc(radius = radius, centre = center)) # Ensure circular window

  # Convert sampled points to ppp object
  points_ppp <- ppp(
    x = points_df$X1,
    y = points_df$Y1,
    window = win
  )
  
  # Define the grid for environmental covariate (angle calculation)
  x_range <- seq(win$xrange[1], win$xrange[2], length.out = 100)  # Increase grid resolution
  y_range <- seq(win$yrange[1], win$yrange[2], length.out = 100)  # Increase grid resolution
  
  # Calculate angle around the circle for each point
  gridcov <- outer(x_range, y_range, function(x, y) {
    # Calculate angle in radians
    atan2(y - center[2], x - center[1])
  })

  # Convert covariate to im object
  cov_im <- im(gridcov, xcol = x_range, yrow = y_range)

  # Adjust the mu to include the time term
  mu_time <- im(
    b0 + b1 * gridcov + b2 * time,
    xcol = x_range,
    yrow = y_range
    )

  lgcp_sim <- rLGCP(
    model = "exp",
    mu = mu_time,
    var = var,
    scale = scale,
    win = win
    )

  # Plot results if requested
  if (plotdat) {
    par(mfrow = c(1, 2), mar = c(2, 2, 1, 1))
    plot(points_ppp, main = "Observed Point Pattern")
    plot(lgcp_sim, main = "Simulated LGCP")
  }

  # Return simulated LGCP and observed points for further analysis
  list(simulated = lgcp_sim, observed = points_ppp)
}

# Parameters
b0 <- -5.5   # Intercept
b1 <- 0.2    # Coefficient for covariate (angle)
b2 <- 1     # Time parameter
time <- 0
var <- 0.1    # Variance
scale <- 50  # Scale

# Analyze each dataset
cat("Analyzing Control dataset...\n")
result_control <- analyze_lgcp(Point_sample_control, circle_center, circle_radius, b0, b1, var, scale)

#Change the time and b2 parameter as required
```
Preventing overlaps
```{r}
# Load required libraries
library(spatstat)
library(dplyr)
library(readxl)
library(lgcp)

# Function to remove overlapping points
remove_overlapping_points <- function(ppp_obj, min_dist) {
  # Iteratively remove points that are too close
  keep <- rep(TRUE, npoints(ppp_obj))
  for (i in seq_len(npoints(ppp_obj))) {
    if (keep[i]) {
      dist_to_others <- nndist(ppp_obj)[i]
      if (dist_to_others < min_dist) {
        keep[-i] <- nndist(ppp_obj[-i]) > min_dist
      }
    }
  }
  subset(ppp_obj, keep)
}

# Function to simulate and analyze LGCP with minimum distance constraint
analyze_lgcp_no_overlap <- function(points_df, center, radius, b0, b1, var, scale, min_dist, plotdat = TRUE) {
  # Define circular window
  win <- owin(
    xrange = c(center[1] - radius, center[1] + radius),
    yrange = c(center[2] - radius, center[2] + radius)
  )
  win <- intersect.owin(win, disc(radius = radius, centre = center)) # Ensure circular window

  # Convert sampled points to ppp object
  points_ppp <- ppp(
    x = points_df$X1,
    y = points_df$Y1,
    window = win
  )
  
  # Define the grid for environmental covariate (angle calculation)
  x_range <- seq(win$xrange[1], win$xrange[2], length.out = 100)
  y_range <- seq(win$yrange[1], win$yrange[2], length.out = 100)
  
  # Calculate angle around the circle for each point
  gridcov <- outer(x_range, y_range, function(x, y) atan2(y - center[2], x - center[1]))
  cov_im <- im(gridcov, xcol = x_range, yrow = y_range)

  # Simulate the LGCP
  lgcp_sim <- rLGCP(
    model = "exp",
    mu = im(b0 + b1 * gridcov, xcol = x_range, yrow = y_range),
    var = var,
    scale = scale,
    win = win
  )
  
  # Apply hard-core thinning to prevent overlapping points
  lgcp_no_overlap <- remove_overlapping_points(lgcp_sim, min_dist)

  # Plot results if requested
  if (plotdat) {
    par(mfrow = c(1, 2), mar = c(2, 2, 1, 1))
    plot(points_ppp, main = "Observed Point Pattern")
    plot(lgcp_no_overlap, main = "LGCP Without Overlap")
  }

  # Return the simulated patterns
  list(simulated = lgcp_sim, no_overlap = lgcp_no_overlap, observed = points_ppp)
}

# Parameters
b0 <- -5.4    # Intercept
b1 <- 0.2     # Coefficient for covariate (angle)
var <- 0.1    # Variance
scale <- 50   # Scale
min_dist <- 3 # Minimum distance to prevent overlaps (units based on data)

# Analyze datasets
cat("Analyzing Control dataset...\n")
result_control <- analyze_lgcp_no_overlap(Point_sample_control, circle_center, circle_radius, b0, b1, var, scale, min_dist)

cat("Analyzing 3-day dataset...\n")
result_3d <- analyze_lgcp_no_overlap(Point_sample_3d, circle_center, circle_radius, b0, b1, var, scale, min_dist)

cat("Analyzing 7-day dataset...\n")
result_7d <- analyze_lgcp_no_overlap(Point_sample_7d, circle_center, circle_radius, b0, b1, var, scale, min_dist)

# Summarize results
cat("Control Dataset:\n")
summary(result_control$no_overlap)
summary(result_control$observed)
cat("3-Day Dataset:\n")
summary(result_3d$no_overlap)
cat("7-Day Dataset:\n")
summary(result_7d$no_overlap)
```
















*Need to reseach which model is better*
Spatial-Temporal LGCP Guassian model
```{r}
# Load required libraries
library(spatstat)
library(sf)
library(sp)
library(raster)
library(fields)
library(rstan)
library(tidyverse)
library(bayesplot)
library(dplyr)

# Function to simulate and analyze spatiotemporal LGCP
analyze_stlgcp <- function(points_df, center, radius, b0, b1, b2, var, scale, times, plotdat = TRUE) {
  # Define circular window
  win <- owin(
    xrange = c(center[1] - radius, center[1] + radius),
    yrange = c(center[2] - radius, center[2] + radius)
  )
  win <- intersect.owin(win, disc(radius = radius, centre = center))  # Ensure circular window

  # Convert sampled points to ppp object without marks (no time-based coloring)
  points_ppp <- ppp(
    x = points_df$X1,
    y = points_df$Y1,
    window = win
  )
  
  # Define grid for spatiotemporal environmental covariates
  x_range <- seq(win$xrange[1], win$xrange[2], length.out = 100)  # Spatial grid resolution
  y_range <- seq(win$yrange[1], win$yrange[2], length.out = 100)  # Spatial grid resolution
  t_range <- seq(min(times), max(times), length.out = 10)         # Temporal grid resolution

  # Generate spatial covariate: angle around the circle
  gridcov_spatial <- outer(x_range, y_range, function(x, y) {
    atan2(y - center[2], x - center[1])  # Angle in radians
  })

  # Simulate the LGCP for each time slice
  lgcp_sims <- list()  # Store simulated LGCPs for each time step

  for (t in t_range) {
    # Compute time-dependent intensity function
    intensity_mu <- im(b0 + b1 * gridcov_spatial + b2 * t, xcol = x_range, yrow = y_range)

    # Simulate LGCP for current time slice
    lgcp_sims[[paste0("t_", t)]] <- rLGCP(
      model = "gauss",     #Need to find out which model to use: one of “exponential”, “gauss”, “stable”, “gencauchy”, “matern”                      
      mu = intensity_mu,
      var = var,
      scale = scale,
      win = win
    )
  }

  # Plot results if requested
  if (plotdat) {
    par(mfrow = c(1, 2), mar = c(2, 2, 1, 1))
    
    # Plot observed point pattern without marks (no color based on time)
    plot(points_ppp, main = "Observed Point Pattern")
    
    # Plot simulated LGCPs
    for (i in seq_along(lgcp_sims)) {
      plot(lgcp_sims[[i]], main = paste("Simulated LGCP at t =", t_range[i]))
    }
  }

  # Return simulated LGCPs and observed points for further analysis
  list(simulated = lgcp_sims, observed = points_ppp)
}

# Parameters
b0 <- -5.5    # Intercept
b1 <- 0.2     # Coefficient for spatial covariate (angle)
b2 <- -0.1    # Coefficient for temporal covariate (time)
var <- 0.1    # Variance
scale <- 50   # Scale
times <- c(0, 3, 7)  # Time points for control, 3-day, and 7-day datasets

# Analyze datasets
cat("Analyzing Control dataset...\n")
result_control <- analyze_stlgcp(Point_sample_control, circle_center, circle_radius, b0, b1, b2, var, scale, times)



# Plot observed points with time as color
plot(result_3d$observed, main = "Observed Points - 3 Day")
plot(result_7d$observed, main = "Observed Points - 7 Day")

# Analyze each dataset
#cat("Analyzing Control dataset...\n")
#result_control <- analyze_stlgcp(Point_sample_control, circle_center, circle_radius, b0, b1, b2, var, scale, times)

#cat("Analyzing 3-day dataset...\n")
#result_3d <- analyze_stlgcp(Point_sample_3d, circle_center, circle_radius, b0, b1, b2, var, scale, times)

#cat("Analyzing 7-day dataset...\n")
#result_7d <- analyze_stlgcp(Point_sample_7d, circle_center, circle_radius, b0, b1, b2, var, scale, times)

# Summaries
#summary(result_control$simulated)
#summary(result_control$observed)

#summary(result_3d$simulated)
#summary(result_3d$observed)

#summary(result_7d$simulated)
#summary(result_7d$observed)
```




Matern Model 
```{r}
# Load required libraries
library(spatstat)
library(sf)
library(sp)
library(raster)
library(fields)
library(rstan)
library(tidyverse)
library(bayesplot)
library(dplyr)

# Function to simulate and analyze spatiotemporal LGCP
analyze_stlgcp <- function(points_df, center, radius, b0, b1, b2, var, scale, times, plotdat = TRUE) {
  # Define circular window
  win <- owin(
    xrange = c(center[1] - radius, center[1] + radius),
    yrange = c(center[2] - radius, center[2] + radius)
  )
  win <- intersect.owin(win, disc(radius = radius, centre = center))  # Ensure circular window

  # Convert sampled points to ppp object without marks (no time-based coloring)
  points_ppp <- ppp(
    x = points_df$X1,
    y = points_df$Y1,
    window = win
  )
  
  # Define grid for spatiotemporal environmental covariates
  x_range <- seq(win$xrange[1], win$xrange[2], length.out = 100)  # Spatial grid resolution
  y_range <- seq(win$yrange[1], win$yrange[2], length.out = 100)  # Spatial grid resolution
  t_range <- seq(min(times), max(times), length.out = 10)         # Temporal grid resolution

  # Generate spatial covariate: angle around the circle
  gridcov_spatial <- outer(x_range, y_range, function(x, y) {
    atan2(y - center[2], x - center[1])  # Angle in radians
  })

  # Simulate the LGCP for each time slice
  lgcp_sims <- list()  # Store simulated LGCPs for each time step

  for (t in t_range) {
    # Compute time-dependent intensity function
    intensity_mu <- im(b0 + b1 * gridcov_spatial + b2 * t, xcol = x_range, yrow = y_range)

    # Simulate LGCP for current time slice using Matern kernel, specifying 'nu'
    lgcp_sims[[paste0("t_", t)]] <- rLGCP(
      model = "matern",     # Using Matern kernel
      mu = intensity_mu,
      var = var,
      scale = scale,
      nu = 1.5,               # Specify nu, e.g., nu = 1 for smoother clusters (Gaussian-like), 1/2= Exp, 1= Gaussian, 1 to 1.5 = more complex clustering
      win = win
    )
}



  # Plot results if requested
  if (plotdat) {
    par(mfrow = c(1, 2), mar = c(2, 2, 1, 1))
    
    # Plot observed point pattern without marks (no color based on time)
    plot(points_ppp, main = "Observed Point Pattern")
    
    # Plot simulated LGCPs
    for (i in seq_along(lgcp_sims)) {
      plot(lgcp_sims[[i]], main = paste("Simulated LGCP at t =", t_range[i]))
    }
  }

  # Return simulated LGCPs and observed points for further analysis
  list(simulated = lgcp_sims, observed = points_ppp)
}

# Parameters
b0 <- -5.5    # Intercept
b1 <- 0.2     # Coefficient for spatial covariate (angle)
b2 <- -0.1    # Coefficient for temporal covariate (time)
var <- 0.1    # Variance
scale <- 50   # Scale
times <- c(0, 3, 7)  # Time points for control, 3-day, and 7-day datasets

# Analyze datasets
cat("Analyzing Control dataset...\n")
result_control <- analyze_stlgcp(Point_sample_control, circle_center, circle_radius, b0, b1, b2, var, scale, times)



plot(result_3d$observed, main = "Observed Points - 3 Day")
plot(result_7d$observed, main = "Observed Points - 7 Day")
```


















Angle and distance from circle as covariates, had to normalize the covariates so finding the parameter values is easier 
```{r}
library(spatstat)
library(sf)
library(sp)
library(raster)
library(fields)
library(rstan)
library(tidyverse)
library(bayesplot)
library(dplyr)

# Function to simulate and analyze spatiotemporal LGCP
analyze_stlgcp <- function(points_df, center, radius, b0, b1, b2, b3, var, scale, times, plotdat = TRUE) {
  # Define circular window
  win <- owin(
    xrange = c(center[1] - radius, center[1] + radius),
    yrange = c(center[2] - radius, center[2] + radius)
  )
  win <- intersect.owin(win, disc(radius = radius, centre = center))  # Ensure circular window

  # Convert sampled points to ppp object without marks (no time-based coloring)
  points_ppp <- ppp(
    x = points_df$X1,
    y = points_df$Y1,
    window = win
  )
  
  # Define grid for spatiotemporal environmental covariates
  x_range <- seq(win$xrange[1], win$xrange[2], length.out = 100)  # Spatial grid resolution
  y_range <- seq(win$yrange[1], win$yrange[2], length.out = 100)  # Spatial grid resolution
  t_range <- seq(min(times), max(times), length.out = 10)         # Temporal grid resolution

  # Generate spatial covariate: angle around the circle
  gridcov_spatial <- outer(x_range, y_range, function(x, y) {
    atan2(y - center[2], x - center[1])/ (2 * pi)  # Angle in radians normalized
  })

  # Generate spatial covariate: distance from the center
  grid_points <- expand.grid(x = x_range, y = y_range)  # Create grid of x, y points
  distances <- sqrt((grid_points$x - center[1])^2 + (grid_points$y - center[2])^2)  # Euclidean distance from the center

  # Convert distances into a matrix form corresponding to the grid
  gridcov_distance_not_normal <- matrix(distances, nrow = length(x_range), ncol = length(y_range), byrow = TRUE)

  # Calculate the minimum and maximum values of the distance matrix for normalization
  min_distance <- min(gridcov_distance_not_normal)
  max_distance <- max(gridcov_distance_not_normal)

  # Normalize the distance covariate
  gridcov_distance <- (gridcov_distance_not_normal - min_distance) / (max_distance - min_distance)  # Normalize

  # Simulate the LGCP for each time slice
  lgcp_sims <- list()  # Store simulated LGCPs for each time step

  for (t in t_range) {
    # Compute time-dependent intensity function, adding the distance covariate with b3
    intensity_mu <- im(b0 + b1 * gridcov_spatial + b2 * t + b3 * gridcov_distance, 
                       xcol = x_range, yrow = y_range)

    # Simulate LGCP for current time slice using Matern kernel, specifying 'nu'
    lgcp_sims[[paste0("t_", t)]] <- rLGCP(
      model = "matern",     # Using Matern kernel
      mu = intensity_mu,
      var = var,
      scale = scale,
      nu = 1.5,               # Specify nu, e.g., nu = 1 for smoother clusters (Gaussian-like), 1/2= Exp, 1= Gaussian, 1 to 1.5 = more complex clustering
      win = win
    )
  }

  # Plot results if requested
  if (plotdat) {
    par(mfrow = c(1, 2), mar = c(2, 2, 1, 1))
    
    # Plot observed point pattern without marks (no color based on time)
    plot(points_ppp, main = "Observed Point Pattern")
    
    # Plot simulated LGCPs
    for (i in seq_along(lgcp_sims)) {
      plot(lgcp_sims[[i]], main = paste("Simulated LGCP at t =", t_range[i]))
    }
  }

  # Return simulated LGCPs and observed points for further analysis
  list(simulated = lgcp_sims, observed = points_ppp)
}

# Parameters
b0 <- -5.5    # Intercept
b1 <- 0.2     # Coefficient for spatial covariate (angle)
b2 <- -0.1    # Coefficient for temporal covariate (time)
b3 <- -0.1     # Coefficient for distance covariate (distance from center)
var <- 0.1    # Variance
scale <- 50   # Scale
times <- c(0, 3, 7)  # Time points for control, 3-day, and 7-day datasets

# Analyze datasets
cat("Analyzing Control dataset...\n")
result_control <- analyze_stlgcp(Point_sample_control, circle_center, circle_radius, b0, b1, b2, b3, var, scale, times)
```


*Still need to find how to choose the correct parameter values*
Add area of points as a mark
```{r}
analyze_stlgcp_with_marks <- function(points_df, center, radius, b0, b1, b2, b3, var, scale, times, plotdat = TRUE) {
  # Define circular window
  win <- owin(
    xrange = c(center[1] - radius, center[1] + radius),
    yrange = c(center[2] - radius, center[2] + radius)
  )
  win <- intersect.owin(win, disc(radius = radius, centre = center))  # Ensure circular window

  # Convert sampled points to ppp object with marks (Area as marks)
  points_ppp <- ppp(
    x = points_df$X1,
    y = points_df$Y1,
    window = win,
    marks = points_df$Area_meters_squared
  )
  
  # Define grid for spatiotemporal environmental covariates
  x_range <- seq(win$xrange[1], win$xrange[2], length.out = 100)  # Spatial grid resolution
  y_range <- seq(win$yrange[1], win$yrange[2], length.out = 100)  # Spatial grid resolution
  t_range <- seq(min(times), max(times), length.out = length(times))  # Temporal grid based on times

  # Generate spatial covariate: angle around the circle
  gridcov_spatial <- outer(x_range, y_range, function(x, y) {
    atan2(y - center[2], x - center[1]) / (2 * pi)  # Angle in radians normalized
  })

  # Generate spatial covariate: distance from the center
  grid_points <- expand.grid(x = x_range, y = y_range)  # Create grid of x, y points
  distances <- sqrt((grid_points$x - center[1])^2 + (grid_points$y - center[2])^2)  # Euclidean distance from the center

  # Convert distances into a matrix form corresponding to the grid
  gridcov_distance_not_normal <- matrix(distances, nrow = length(x_range), ncol = length(y_range), byrow = TRUE)

  # Calculate the minimum and maximum values of the distance matrix for normalization
  min_distance <- min(gridcov_distance_not_normal)
  max_distance <- max(gridcov_distance_not_normal)

  # Normalize the distance covariate
  gridcov_distance <- (gridcov_distance_not_normal - min_distance) / (max_distance - min_distance)  # Normalize

  # Simulate the LGCP for each time slice
  lgcp_sims <- list()  # Store simulated LGCPs for each time step

  for (t in t_range) {
    # Compute time-dependent intensity function, adding the distance covariate with b3
    intensity_mu <- im(b0 + b1 * gridcov_spatial + b2 * t + b3 * gridcov_distance, 
                       xcol = x_range, yrow = y_range)

    # Simulate LGCP for current time slice using Matern kernel
    lgcp_sims[[paste0("t_", t)]] <- rLGCP(
      model = "matern",     # Using Matern kernel
      mu = intensity_mu,
      var = var,
      scale = scale,
      nu = 1.5,               # Specify nu
      win = win
    )
  }

  # Assign marks to simulated LGCP points
  simulate_marks <- function(simulated_ppp, observed_marks) {
    # Assign marks sampled from the observed marks distribution
    simulated_ppp$marks <- sample(observed_marks, size = simulated_ppp$n, replace = TRUE)
    return(simulated_ppp)
  }

  # Apply marks to all simulated LGCP points
  lgcp_sims_with_marks <- lapply(lgcp_sims, simulate_marks, observed_marks = points_df$Area_meters_squared)

  # Plot results if requested
  if (plotdat) {
    n_plots <- length(times) + 1  # Number of plots: one for observed + one for each time step
    n_cols <- 2                  # Number of columns in the plot layout
    n_rows <- ceiling(n_plots / n_cols)  # Calculate number of rows for the layout
    par(mfrow = c(n_rows, n_cols), mar = c(2, 2, 1, 1))
    
    # Plot observed point pattern with marks (Area)
    plot(points_ppp, main = "Observed Point Pattern")
    
    # Plot simulated LGCPs for each time slice
    for (i in seq_along(times)) {
      plot(lgcp_sims_with_marks[[i]], main = paste("Simulated LGCP at t =", times[i]))
    }
  }

  # Return simulated LGCPs and observed points for further analysis
  list(simulated = lgcp_sims_with_marks, observed = points_ppp)
}

# Parameters
b0 <- -5.4    # Intercept
b1 <- 0.1     # Coefficient for spatial covariate (angle), too high, it may create strong clustering around specific angular regions. Making this value large causes clustering at certain                   angles
b2 <- -0.1    # Coefficient for temporal covariate (time), too high, it could amplify clustering based on the time slices.
b3 <- 0.1    # Coefficient for distance covariate (distance from center), too high, it may create strong clustering around specific distances. Making this value large moves everything away                from the centre
var <- 0.05    # Variance, lower value smoother intensity function
scale <- 100   # Scale, increase to spread out the influence of individual points
times <- c(0, 3, 7)  # Time points for control, 3-day, and 7-day datasets

# Analyze the datasets
cat("Analyzing Control dataset...\n")
result_control <- analyze_stlgcp_with_marks(Point_sample_control, circle_center, circle_radius, b0, b1, b2, b3, var, scale, times)

cat("Analyzing 3-day dataset...\n")
result_3d <- analyze_stlgcp_with_marks(Point_sample_3d, circle_center, circle_radius, b0, b1, b2, b3, var, scale, times)

cat("Analyzing 7-day dataset...\n")
result_7d <- analyze_stlgcp_with_marks(Point_sample_7d, circle_center, circle_radius, b0, b1, b2, b3, var, scale, times)
```



Density plot and area of point pattern is included in the calculation
```{r}
library(spatstat)
library(sf)
library(sp)
library(raster)
library(fields)
library(rstan)
library(tidyverse)
library(bayesplot)
library(dplyr)

analyze_stlgcp_with_marks <- function(points_df, center, radius, b0, b1, b2, b3, var, scale, times, plotdat = TRUE) {
  # Define circular window
  win <- owin(
    xrange = c(center[1] - radius, center[1] + radius),
    yrange = c(center[2] - radius, center[2] + radius)
  )
  win <- intersect.owin(win, disc(radius = radius, centre = center))  # Ensure circular window

  # Convert sampled points to ppp object with marks (Area as marks)
  points_ppp <- ppp(
    x = points_df$X1,
    y = points_df$Y1,
    window = win,
    marks = points_df$Area_meters_squared
  )
  
  # Define grid for spatiotemporal environmental covariates
  x_range <- seq(win$xrange[1], win$xrange[2], length.out = 100)
  y_range <- seq(win$yrange[1], win$yrange[2], length.out = 100)
  t_range <- seq(min(times), max(times), length.out = length(times))

  # Generate spatial covariates
  gridcov_spatial <- outer(x_range, y_range, function(x, y) {
    atan2(y - center[2], x - center[1]) / (2 * pi)
  })

  grid_points <- expand.grid(x = x_range, y = y_range)
  distances <- sqrt((grid_points$x - center[1])^2 + (grid_points$y - center[2])^2)
  gridcov_distance <- matrix(
    (distances - min(distances)) / (max(distances) - min(distances)),
    nrow = length(x_range), ncol = length(y_range), byrow = TRUE
  )

  # Simulate LGCP for each time slice
  lgcp_sims <- list()
  for (t in t_range) {
    intensity_mu <- im(b0 + b1 * gridcov_spatial + b2 * t + b3 * gridcov_distance, xcol = x_range, yrow = y_range)
    lgcp_sims[[paste0("t_", t)]] <- rLGCP(
      model = "matern",
      mu = intensity_mu,
      var = var,
      scale = scale,
      nu = 1.5,
      win = win
    )
  }

  # Assign marks to simulated LGCP points
  simulate_marks <- function(simulated_ppp, observed_marks) {
    simulated_ppp$marks <- sample(observed_marks, size = simulated_ppp$n, replace = TRUE)
    return(simulated_ppp)
  }
  lgcp_sims_with_marks <- lapply(lgcp_sims, simulate_marks, observed_marks = points_df$Area_meters_squared)

  # Plot results if requested
  if (plotdat) {
    # Calculate densities considering marks (Area_meters_squared)
    observed_density <- density(
      points_ppp,
      weights = points_ppp$marks,  # Use the area as weights
      sigma = bw.diggle(points_ppp),
      eps = 0.1
    )
    simulated_densities <- lapply(lgcp_sims_with_marks, function(ppp) {
      density(
        ppp,
        weights = ppp$marks,  # Use simulated marks
        sigma = bw.diggle(ppp),
        eps = 0.2
      )
    })

    # Find global zlim (shared scale across all densities)
    all_density_values <- c(observed_density$v, unlist(lapply(simulated_densities, function(d) d$v)))
    global_min <- min(all_density_values, na.rm = TRUE)
    global_max <- max(all_density_values, na.rm = TRUE)

    # Ensure zlim is valid
    if (global_min >= global_max) {
      stop("Invalid zlim: Ensure density values are not degenerate.")
    }

    # Plot observed and simulated densities
    n_plots <- length(times) + 1
    n_cols <- 2
    n_rows <- ceiling(n_plots / n_cols)
    par(mfrow = c(n_rows, n_cols), mar = c(2, 2, 1, 1))
    
    plot(observed_density, main = "Observed Point Density", zlim = c(global_min, global_max), ribscale = 1.5, ribargs = list(las = 1))
    for (i in seq_along(times)) {
      plot(simulated_densities[[i]], main = paste("Simulated Density at t =", times[i]), zlim = c(global_min, global_max), ribscale = 1.5, ribargs = list(las = 1))
    }
  }

  list(simulated = lgcp_sims_with_marks, observed = points_ppp)
}

# Parameters
b0 <- -5.8    # Intercept
b1 <- 0.1     # Coefficient for spatial covariate (angle), too high, it may create strong clustering around specific angular regions. Making this value large causes clustering at certain                   angles
b2 <- -0.12    # Coefficient for temporal covariate (time), too high, it could amplify clustering based on the time slices.
b3 <- 0.1    # Coefficient for distance covariate (distance from center), too high, it may create strong clustering around specific distances. Making this value large moves everything away                from the centre
var <- 0.3    # Variance, lower value smoother intensity function
scale <- 0.2   # Scale, increase to spread out the influence of individual points
times <- c(0, 3, 7)  # Time points for control, 3-day, and 7-day datasets
 
# Analyze the datasets
cat("Analyzing Control dataset...\n")
result_control <- analyze_stlgcp_with_marks(Point_sample_control, circle_center, circle_radius, b0, b1, b2, b3, var, scale, times)

cat("Analyzing 3-day dataset...\n")
result_3d <- analyze_stlgcp_with_marks(Point_sample_3d, circle_center, circle_radius, b0, b1, b2, b3, var, scale, times)

cat("Analyzing 7-day dataset...\n")
result_7d <- analyze_stlgcp_with_marks(Point_sample_7d, circle_center, circle_radius, b0, b1, b2, b3, var, scale, times)
```

Minimum Distance from centre 
Points removed if they are not minimum distance
```{r}

# Function to enforce minimum distance between points
enforce_min_distance <- function(points_ppp, min_dist) {
  # Use spatstat's `rthin` to thin points based on the minimum distance
  while (TRUE) {
    # Compute the pairwise distances between points
    pair_dists <- pairdist(points_ppp)
    # Set the diagonal to infinity to exclude self-distances
    diag(pair_dists) <- Inf
    # Check if all pairwise distances are greater than the minimum distance
    if (all(pair_dists > min_dist)) break
    # Remove a random point that violates the minimum distance
    points_ppp <- points_ppp[-sample(1:points_ppp$n, 1)]
  }
  return(points_ppp)
}

analyze_stlgcp_with_marks <- function(points_df, center, radius, b0, b1, b2, b3, var, scale, times, min_dist = NULL, plotdat = TRUE) {
  # Define circular window
  win <- owin(
    xrange = c(center[1] - radius, center[1] + radius),
    yrange = c(center[2] - radius, center[2] + radius)
  )
  win <- intersect.owin(win, disc(radius = radius, centre = center))  # Ensure circular window

  # Convert sampled points to ppp object with marks (Area as marks)
  points_ppp <- ppp(
    x = points_df$X1,
    y = points_df$Y1,
    window = win,
    marks = points_df$Area_meters_squared
  )
  
  # Define grid for spatiotemporal environmental covariates
  x_range <- seq(win$xrange[1], win$xrange[2], length.out = 100)
  y_range <- seq(win$yrange[1], win$yrange[2], length.out = 100)
  t_range <- seq(min(times), max(times), length.out = length(times))

  # Generate spatial covariates
  gridcov_spatial <- outer(x_range, y_range, function(x, y) {
    atan2(y - center[2], x - center[1]) / (2 * pi)
  })
  grid_points <- expand.grid(x = x_range, y = y_range)
  distances <- sqrt((grid_points$x - center[1])^2 + (grid_points$y - center[2])^2)
  gridcov_distance <- matrix(distances, nrow = length(x_range), byrow = TRUE)
  gridcov_distance <- (gridcov_distance - min(gridcov_distance)) / diff(range(gridcov_distance))

  # Simulate LGCP for each time slice
  lgcp_sims <- list()
  for (t in t_range) {
    intensity_mu <- im(b0 + b1 * gridcov_spatial + b2 * t + b3 * gridcov_distance, xcol = x_range, yrow = y_range)
    lgcp_sim <- rLGCP("matern", mu = intensity_mu, var = var, scale = scale, nu = 1.5, win = win)
    
    # Enforce minimum distance if specified
    if (!is.null(min_dist)) {
      lgcp_sim <- enforce_min_distance(lgcp_sim, min_dist)
    }
    
    lgcp_sims[[paste0("t_", t)]] <- lgcp_sim
  }

  # Assign marks to simulated points
  simulate_marks <- function(simulated_ppp, observed_marks) {
    simulated_ppp$marks <- sample(observed_marks, size = simulated_ppp$n, replace = TRUE)
    return(simulated_ppp)
  }
  lgcp_sims_with_marks <- lapply(lgcp_sims, simulate_marks, observed_marks = points_df$Area_meters_squared)

  # Plot results
  if (plotdat) {
    n_plots <- length(times) + 1
    par(mfrow = c(ceiling(n_plots / 2), 2), mar = c(2, 2, 1, 1))
    plot(points_ppp, main = "Observed Point Pattern")
    for (i in seq_along(times)) {
      plot(lgcp_sims_with_marks[[i]], main = paste("Simulated LGCP at t =", times[i]))
    }
  }

  # Return results
  list(simulated = lgcp_sims_with_marks, observed = points_ppp)
}


# Parameters
b0 <- -5.8    # Intercept
b1 <- 0.1     # Coefficient for spatial covariate (angle), too high, it may create strong clustering around specific angular regions. Making this value large causes clustering at certain                   angles
b2 <- -0.12    # Coefficient for temporal covariate (time), too high, it could amplify clustering based on the time slices.
b3 <- 0.1    # Coefficient for distance covariate (distance from center), too high, it may create strong clustering around specific distances. Making this value large moves everything away                from the centre
var <- 0.3    # Variance, lower value smoother intensity function
scale <- 0.2   # Scale, increase to spread out the influence of individual points
times <- c(0, 3, 7)  # Time points for control, 3-day, and 7-day datasets
min_dist = 0.005


# Analyze the datasets
cat("Analyzing Control dataset...\n")
result_control <- analyze_stlgcp_with_marks(Point_sample_control, circle_center, circle_radius, b0, b1, b2, b3, var, scale, times, min_dist = 0.0005)

```